---
title: "Gov 50 Final Project"
author: "Ethan Johnstone"
description: "Trump Campaign Stops impact on 2016 Voter Turnout"
output:
  distill::distill_article:
    self_contained: false
---

## Proposal

Do campaign stops impact voter turnout in Presidential elections? In this study, I plan to examine the relationship between Donald Trump's 2016 campaign stops and voter turnout at the county level. I hypothesize that Donald Trump's campaign stops increased overall voter turnout compared to counties that he did not visit. It is widely held that Trump used large campaign stops to garner energy, funds, and turnout voters in rural counties in swing states such as Wisconsin, Michigan, and Pennsylvania. This study will evaluate if this theory can be substantiated with voter turnout data from the 2016 election. My sample is comprised of county level voter turnout data compiled by the MIT Election Lab. My unit of analyses is Trump camping stops. My explanatory variable of interest is whether Trump held a campaign stop in a county or not. The variable is coded =1 if he held a campaign stop and =0 if he did not. My outcome variable is the percent of eligible voters that cast votes. For instance if this variable was coded =56, then this would indicate that 56% of eligible voters voted in that county. If I observe higher voter turnout in counties that Trump held campaign stops in, the data would support my hypothesis. If voter turnout was unchanged or less in counties he held stops in, then the data would disprove my hypothesis.

## Introduction

Do campaign stops impact voter turnout in Presidential elections? In this study, I plan to examine the relationship between Donald Trump's 2016 campaign stops and voter turnout at the county level. I hypothesize that Donald Trump's campaign stops increased overall voter turnout compared to counties that he did not visit. It is widely held that Trump used large and entertaining campaign stops to garner energy, funds, and turnout voters in rural counties in swing states such as Wisconsin, Michigan, and Pennsylvania. This study will evaluate if this theory can be substantiated with voter turnout data from the 2016 election. In the field of Political Science and in the world of US politics there has long been an argument of whether or not campaigning even has an impact on presidential elections. Some political scientists argue that election outcomes are more based on political environments and partisan pre-disposition than on the efforts of campaigns. The findings in this study will hopefully help us understand the relationship between campaign stops and voter turnout -- which can change the outcome of an election in close elections such as in 2016. If the data suggests that campaign visits in 2016 increased voter turnout, then this would lend support to the notion that campaigns matter for the outcome of elections.

## Data

```{r}
library(tidyverse)
library(readxl)
library(ggplot2)
library(dbplyr)
library("openair")
rally <- read.csv("Trump_Rally_County_1.csv")
turnout <- read_xlsx("Data/Turnout_Data_2016.xlsx", sheet = "County VTO")
names(turnout)[1:2] <- c("countyname", "state")
names(turnout)[11] <- c("turnout_perc")
names(rally)[3] <- c("countyname")

final_data <- left_join(turnout, rally, by = c("countyname"))

whole_data <- final_data |> select(c(1,2,3,9,11,81,82,83))

whole_data[["Rally"]][is.na(whole_data[["Rally"]])] <- 0

whole_data <- whole_data |> mutate(Rally_ph = if_else(Rally == 1, "Event", "No Event"))
tibble(whole_data)

```

My sample is comprised of county level voter turnout data compiled by the MIT Election Lab from the 2016 election. My unit of analysis is official Trump campaign stops. My explanatory variable of interest is whether Trump held a campaign stop in a county or not. The variable is coded =1 if he held a campaign stop and =0 if he did not. This independent variable is measured in expert coding and is from a list that was compiled by staffers in the 2016 Trump campaign. My outcome variable is the percent of eligible voters that cast votes. For instance if this variable was coded =56, then this would indicate that 56% of eligible voters voted in that county. If I observe higher voter turnout in counties that Trump held campaign stops in, the data would support my hypothesis. If voter turnout was unchanged or less in counties he held stops in, then the data would disprove my hypothesis.

Distribution of Dependent Variable:
```{r}
turnout_dist <- ggplot(whole_data, aes(x=turnout_perc)) +
  labs(title = "Distribution of Voter Turnout",
       x = "Voter Turnout Percentage",
       y = "Number of Counties") +
  geom_histogram()
turnout_dist

```

## Results

```{r}
ggplot(data = whole_data,
       mapping = aes(group = Rally_ph,
                     x = Rally_ph,
                     y = turnout_perc)) +
labs(title = "County Turnout and Trump Campaign Events",
     x = "Campaign Event",
     y = "Turnout Percentage") +
  geom_boxplot()

turnout_regression <- lm(turnout_perc ~ Rally, data = whole_data)
modelsummary::modelsummary(turnout_regression,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"))

```

This study is a cross sectional study. The coefficient for 'Rally,' which is the main coefficient of interest is .007, which means that there is a very slight positive correlation between Trump campaign events and voter turnout at the county level. However, this correlation is not statistically significant since the p-value is .104 and falls well outside of the acceptable range of less than .05. I do not believe the coefficient represents a casual effect since the correlation is too small, is not statistically significant, and as the box plot demonstrates there is not a significant variance in voter turnout among counties that had a Trump campaign stop versus those that did not.


## Summary of Findings






